Quantifying Lewontin's Paradox Suggests Natural Selection is Unlikely to Explain the Narrow Range of Diversity Among Species

### 2021-06-04 manuscript/elife/elife-template.tex
more elife template bullshit

\documentclass[11pt]{article}
\RequirePackage{fullpage}
\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{mathtools}
\RequirePackage{graphicx}
\RequirePackage[normalem]{ulem}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{wrapfig}
\RequirePackage{subcaption}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
\RequirePackage{tikz}
\RequirePackage{wrapfig}

% line numbers:
\RequirePackage{lineno}
% \modulolinenumbers[5]
\linenumbers
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{black}}

% spacing
\RequirePackage{setspace}
% \doublespacing

% \RequirePackage[osf]{mathpazo}
\RequirePackage[bibstyle=authoryear,citestyle=authoryear-comp,
                date=year,
                maxbibnames=9,maxnames=5,maxcitenames=2,
                backend=biber,uniquelist=false,uniquename=false,
                % style=apa,
                sorting=nyt,
                % sorting=,
                hyperref=true]{biblatex}
\RequirePackage[colorinlistoftodos]{todonotes}  %disable
\RequirePackage{color}
\RequirePackage{nicefrac}

\newcommand{\gc}[1]{{\it \color{red} #1 } }
\newcommand{\vb}[1]{{\it \color{blue} #1}}
\newcommand{\vbout}[1]{{\it \color{blue} \sout{#1}}}

% a /nonumber you can turn on/off
\newcommand{\nnn}{\nonumber}
%\newcommand{\nnn}{}

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\newcommand{\Newnameref}[1]{\textit{\nameref{#1}}}

\renewbibmacro{in:}{}
\AtEveryBibitem{\clearlist{language}}

\newcommand{\graham}[1]{\todo[size=\scriptsize, color=red!50]{#1}}
\newcommand{\vince}[1]{\todo[size=\scriptsize, color=blue!50]{#1}}

@@

\newcommand{\chapquote}[2]{\begin{quotation} \textit{#1} \end{quotation} \begin{flushright} - #2\end{flushright} }

\addbibresource{biblio.bib}

\title{Natural Selection is Unlikely to Explain Why Species Get a Thin Slice of $\pi$}
%\title{Why do species get a thin slice of $\pi$? Revisiting Lewontin's Paradox of Variation}

\author{Vince Buffalo}
\affil[]{\footnotesize University of Oregon, Institute of Ecology and Evolution\\Eugene, Oregon \\ \href{mailto:vsbuffalo@gmail.com}{vsbuffalo@gmail.com}}

\begin{document}
\maketitle

% TODO recheck node height and other p-values

% Neutral theory predicts that genetic diversity increases with population size,
% yet observed levels of diversity across species vary only two orders of
% magnitude while population sizes likely vary over several. The causes of this
% discrepancy, known as Lewontin's Paradox of Variation, remain unknown. Here I
% revisit Lewontin's Paradox by quantifying the relationship between pairwise
% diversity and approximate census size for 172 metazoan species. Using
% phylogenetic comparative methods, I show this relationship is significant
% accounting for phylogeny, but has high phylogenetic signal and some lineages
% experience shifts in diversity deep in the past. I find a negative relationship
% between recombination map length and census size, suggesting abundant species
% have less recombination and could experience greater reductions in diversity
% due to linked selection. However, I show that even using strong selection
% parameter estimates, models of linked selection are unlikely to explain the
% observed relationship between diversity and census sizes across species.

@@

 Neutral theory predicts that genetic diversity increases with population size,
 yet observed levels of diversity across metazoans vary only two orders of
 magnitude while population sizes vary over several. This unexpectedly narrow
 range of diversity is known as Lewontin's Paradox of Variation (1974). While
 some have suggested selection constrains diversity, tests of this hypothesis
 seem to fall short. Here, I revisit Lewontin's Paradox to assess whether
 current models of linked selection are capable of reducing diversity to this
 extent. To quantify the discrepancy between pairwise diversity and census
 population sizes across species, I combine previously-published estimates of
 pairwise diversity from 172 metazoan taxa with estimates of census sizes.
 Using phylogenetic comparative methods, I show this relationship is
 significant accounting for phylogeny, but with high phylogenetic signal and
 evidence that some lineages experience shifts in the evolutionary rate of
 diversity deep in the past. Additionally, I find a negative relationship
 between recombination map length and census size, suggesting abundant species
 have less recombination and experience greater reductions in diversity due to
 linked selection. However, I show that even using strong selection parameter
 estimates, models of linked selection are unlikely to explain the observed
 relationship between diversity and census sizes across species.

\end{abstract}

A longstanding mystery in evolutionary genetics is that the observed levels of
genetic variation across sexual species are confined to an unexpectedly narrow
range. Under neutral theory, the average number of nucleotide differences
between lineages (pairwise diversity, $\pi$) is determined by the balance of
new mutations and their loss by genetic drift
\parencite{Kimura1964-ul,Malecot1948-zv,Wright1931-fl}. In particular, the
expected diversity at neutral sites in a panmictic population of $N_c$ diploids
is expected to be $\pi \approx 4N_c \mu$, where $\mu$ is the per basepair per
generation mutation rate. Given that metazoan germline mutation rates only
differ 10-fold ($10^{-8}$--$10^{-9}$, \cite{Kondrashov2010-fi,Lynch2010-ki}),
and census sizes vary over several orders of magnitude, one would expected
under neutral theory that heterozygosity should also vary over several orders
of magnitude. However, early allozyme surveys revealed that heterozygosity
levels across a wide range of species varied just an order of magnitude
\parencite[][p.  208]{Lewontin1974-jb}; this is known as Lewontin's ``Paradox
of Variation".  With modern sequencing-based estimates of $\pi$ across taxa
ranging over only three orders of magnitude ($0.01$--$10\%$,
\cite{Leffler2012-zj}), Lewontin's paradox has persisted unresolved through the
genomics era.

From the beginning, explanations for Lewontin's Paradox have been framed in
terms of the neutralist--selectionist controversy
\parencite{Lewontin1974-jb,Kimura1984-ia,Gillespie1991-qa,Gillespie2001-mv}.
The neutralist view is that beneficial alleles are sufficiently rare and
deleterious alleles removed sufficiently quickly, that levels of genetic
diversity are shaped predominantly by genetic drift and mutation
\parencite{Kimura1984-ia}. Specifically, \emph{non-selective} processes
decouple the effective population size implied by observed levels of diversity
$\widehat{\pi}$, $\widetilde{N}_e = \nicefrac{\widehat{\pi}}{4\mu}$, from the
census size, $N_c$. By contrast, the selectionist view is that the direct and
indirect effects of linked selection suppress diversity levels across taxa,
specifically because the impact of linked selection is greater in large
populations. Undoubtedly, these opposing views represent a false dichotomy, as
population genomic studies have uncovered both complex demographic histories
that impact diversity within a species (e.g.
\cite{Zhao2013-vd,Palkopoulou2015-bg}) and evidence that selection depresses
genome-wide diversity (e.g.
\cite{Elyashiv2016-vt,Begun1992-ey,Aguade1989-jx,McVicker2009-ax}).

\subsection*{Possible Explanations of Lewontin's Paradox}

A resolution of Lewontin's Paradox would involve a mechanistic description and
quantification of the evolutionary processes that prevent diversity from
scaling with census sizes across species. This would necessarily connect to the
broader literature on the empirical relationship between diversity and
population size
\parencite{Frankham1996-yb,Nei1984-zi,Soule1976-he,Leroy2021-gy}, and the
ecological and life history correlates of genetic diversity
\parencite{Nevo1978-wh,Powell1975-lg,Nevo1984-hp}. Three categories of
processes stand out as potentially capable of decoupling census sizes from
diversity: non-equilibrium demography, variance and skew in reproductive
success, and selective processes.

It has long been appreciated that effective population sizes are typically less
than census population sizes, tracing back to early debates between R.A. Fisher
and Sewall Wright \parencite{Fisher1947-tf,Wright1948-fj}. Possible causes of
this divergence between effective and census population sizes include
demographic history (e.g. population bottlenecks), extinction and
recolonization dynamics, or the breeding structure of populations (e.g. the
variance in reproductive success and population substructure). Early
explanations for Lewontin's Paradox suggested bottlenecks during the last
glacial maximum severely reduced population sizes
\parencite{Kimura1984-ia,Ohta1973-pk,Nei1984-zi}, and emphasized that large
populations recover to equilibrium diversity levels more slowly
(\cite{Nei1984-zi}, \cite{Kimura1984-ia} p. 203-204). Another explanation is
that cosmopolitan species repeatedly endure extinction and recolonization
events, which reduces effective population size
\parencite{Maruyama1980-xz,Slatkin1977-kd}.

While chance demographic events like bottlenecks and recent expansions have
long-term impacts on diversity (since mutation-drift equilibrium is reached on
the order of size of the population), characteristics of the breeding structure
such as high variance ($V_w$) or skew in reproductive success continuously
suppress diversity below the levels predicted by the census size
\parencite{Wright1938-tv}. For example, in many marine animals, females are
highly fecund, and dispersing larvae face extremely low survivorship, leading
to high variance in reproductive success
\parencite{Waples2018-kb,Waples2013-wi,Hedgecock2011-ku,Hauser2008-fd}. Such
``sweepstakes" reproductive systems can lead to remarkably small ratios of
effective population size to census population size (e.g. $\nicefrac{N_e}{N_c}$
can range from $10^{-6}$--$10^{-2}$), since $\nicefrac{N_e}{N} \approx
\nicefrac{1}{V_w}$
\parencite{Hedgecock1994-gs,Wright1938-tv,Nunney1993-ef,Nunney1996-wy}, and
require multiple-merger coalescent processes to describe their genealogies
\parencite{Eldon2006-ui}. Overall, these reproductive systems diminish the
diversity in many species, but seem unlikely to explain Lewontin's Paradox
broadly across metazoans.

Alternatively, selective processes, and in particular the indirect effects of
selection on linked neutral variation, could explain the observed narrow range
of diversity. The earliest mathematical model of hitchhiking was proffered as a
explanation of Lewontin's Paradox \parencite{Maynard_Smith1974-zr}.  Since,
empirical observations have demonstrated that linked selection shapes patterns
of genome-wide diversity, as evidenced by the correlation between recombination
and diversity in a variety of species
\parencite{Aguade1989-jx,Begun1992-ey,Cutter2003-tl,Stephan1998-hh,Cai2009-by}.
Theoretic work to explain this pattern has considered diversity under a steady
influx of new beneficial mutations (recurrent hitchhiking;
\cite{Stephan1992-jc,Stephan1995-ry}), and purifying selection against new
deleterious mutations (background selection, BGS;
\cite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1994-oh,Hudson1995-xc}).
Indeed, empirical work indicates background selection diminishes diversity
around genic regions in a variety of species
\parencite{McVicker2009-ax,Hernandez2011-gs,Charlesworth1996-px}, and now
efforts have shifted towards teasing apart the effects of positive and negative
selection on genomic diversity \parencite{Elyashiv2016-vt}.

An important class of theoretic selection models pertaining to Lewontin's
Paradox are recurrent hitchhiking models that decouple diversity from the
census population size. These models predict diversity when strongly selected
beneficial mutations regularly enter and sweep through the population, trapping
lineages and forcing them to coalesce
\parencite{Kaplan1989-sc,Gillespie2000-mh}. In general, decoupling occurs under
these hitchhiking models when the rate of coalescence due to selection is much
greater than the rate of neutral coalescence \parencite[e.g.][equation
22]{Coop2012-cd}. Under other linked selection models, the resulting effective
population size is proportional to population size, and thus these models
cannot decouple diversity, all else equal. For example, models of background
selection and polygenic fitness variation predict diversity is proportional to
population size, mediated by the total recombination map length and the
deleterious mutation rate or fitness variation
\parencite{Charlesworth1993-gb,Nicolaisen2012-vs,Nordborg1996-nq,Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}.

\subsection*{Recent Approaches Towards Solving Lewontin's Paradox}

Recently, \textcite{Corbett-Detig2015-gt} used population genomic data to
estimate the reduction in diversity due to background selection and hitchhiking
across 40 species, and showed the impact of selection increases with two
proxies of census population size, species range and with body size. They
argued this is evidence that selection could explain Lewontin's Paradox;
however, in a re-analysis, \textcite{Coop2016-gx} demonstrated that the
observed scale of these reductions is insufficient to explain the
orders-of-magnitude shortfall between observed and expected levels of diversity
across species. Other recent work has found that certain life history
characteristics related to parental investment, such as propagule size, are
good predictors diversity in animals \parencite{Romiguier2014-bp,Chen2017-nf}.
Nevertheless, while these diversity correlates are important clues, they do not
propose a mechanism by which these traits act to constrain diversity within a
few orders of magnitude.

@@

Past work looking at the relationship between $\pi$ and $N_c$ has been unable
to fully account for phylogenetic non-independence across taxa
\parencite{Felsenstein1985-an}. To address this shortcoming, I use phylogenetic
comparative methods (PCMs) with a synthetic time-calibrated phylogeny to
account for shared phylogenetic history. Moreover, it is disputed whether
considering phylogenetic non-independence is necessary in population genetics,
since coalescent times are much less than divergence times
\parencite{Whitney2010-ud,Lynch2011-qv}. Using PCMs, I address this by
estimating the degree of phylogenetic signal in the diversity census size
relationship, and investigating how these traits evolve along the phylogeny.

Finally, I explore whether the predicted reductions of diversity under
background selection and recurrent hitchhiking are sufficiently strong to
resolve Lewontin's Paradox. These predicted reductions in diversity across
species are estimated using strong selection parameters from \emph{Drosophila
melanogaster}, a species known to be strongly affected by linked selection.
Given the effects of linked selection are mediated by recombination map length,
I investigate how recombination map lengths vary with census population size
using data from a previously-published survey \parencite{Stapley2017-fs}. I
find map lengths are typically shorter in large--census-size species,
increasing the effects of linked selection in these species, which might
further decouple diversity from census size. Still, I find the combined impact
of these selection models with available parameter estimates falls short in
explaining Lewontin's Paradox, and discuss future avenues through which the
Paradox of Variation could be fully resolved.

\section*{Results}

\subsection*{Estimates of Census Population Size}

\begin{figure}[t!]
  \centering
  \includegraphics[width=0.6\textwidth]{figures/Nc_figure.pdf}
  \caption{
     The distribution of approximate census population sizes estimated by this
     study. Some phyla containing few species were excluded for clarity.}
\label{fig:figure-nc}
\end{figure}

A major impediment in quantifying Lewontin's Paradox has been estimating census
population sizes across many taxa, especially for extremely abundant,
cosmopolitan species that define the upper limit of ranges. Previous work has
surveyed the literature for census size estimates
\parencite{Nei1984-zi,Soule1976-he,Frankham1996-yb}, or used range, body size,
or qualitative categories as proxies for census size
\parencite{Corbett-Detig2015-gt,Leffler2012-zj}. To quantify the relationship
between genomic estimates of diversity and census population sizes, I first
approximate census population sizes for 172 metazoan taxa (Figure
\ref{fig:figure-nc}). My approach predicts population densities from body sizes
using a previously-observed linear relationship that holds across metazoans
(Figure 1-figure supplement 1 \ref{suppfig:damuth};
\cite{Damuth1981-st,Damuth1987-sg}). Then, from geographic occurrence data, I
estimate range sizes.  Finally, I estimate population size as the product of
these predicted densities and range estimates (see Methods:
\Newnameref{sec:methods-popsize}). Note that the relationship between
population density and body size is driven by energy budgets, and thus reflects
macroecological equilibria \parencite{Damuth1987-sg}. Consequently, population
sizes are underestimated for taxa like humans and their domesticated species,
and overestimated for species with anthropogenically reduced densities or
fragmented ranges. For example, the population size of \emph{Lynx lynx} is
likely around 50,000 \parencite{Iucn2020-ap} which is around two orders of
magnitude smaller than my estimate. Additionally, the range size estimates do
not consider whether an area has unsuitable habitat, and thus may be
overestimated for species with particular niches or patchy habitats. While
these methods to estimate census size are crude and approximate, they can be
efficiently calculated for numerous taxa and are sufficient to estimate the
scale of Lewontin's Paradox (see Supplementary Information
\Newnameref{supp:popsize-valid} for more on validation based on biomass and
other approaches).

\subsection*{Quantifying Lewontin's Paradox}

\begin{figure}[t!]
  \centering
  \includegraphics[width=\textwidth]{figures/diversity_popsize_full.pdf}

  \caption{An illustration of Lewontin's Paradox of Variation. Pairwise
    diversity (data from \cite{Leffler2012-zj}, \cite{Corbett-Detig2015-gt},
    and \cite{Romiguier2014-bp}), which varies around three orders of
    magnitude, shows a weak relationship with approximate population size,
    which varies over 12 orders of magnitude.  The shaded curve shows the range
    of expected neutral diversity if $N_e$ were to equal $N_c$ under the
    four-alleles model, $\log_{10}(\pi) = \logt(\theta) - \logt{(1 +
    \nicefrac{4\theta}{3})}$ where $\theta = 4N_c \mu$, for two mutation rates,
    $\mu = 10^{-8}$ and $\mu = 10^{-9}$, and the light gray dashed line
    represents the maximum pairwise diversity under the four alleles model. The
    dark gray dashed line is the OLS regression fit, and the blue dashed line
    is the regression fit using a phylogenetic mixed-effects model.  Points are
    colored by phylum. The species \emph{Equus ferus przewalskii} ($N_c \approx
  10^3$ and $\pi = 3.6 \times 10^{-3}$) was an outlier and excluded from this
figure for visual clarity.}

  \label{fig:figure-1}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ELIFE ARTICLE TEMPLATE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PREAMBLE 
\documentclass[9pt,lineno]{elife}
% Use the onehalfspacing option for 1.5 line spacing
% Use the doublespacing option for 2.0 line spacing
% Please note that these options may affect formatting.
% Additionally, the use of the \newcommand function should be limited.

\usepackage{lipsum} % Required to insert dummy text
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\usepackage{nicefrac}
\usepackage{mathtools}
\DeclareSIUnit\Molar{M}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Quantifying Lewontin's Paradox Suggests Natural Selection is Unlikely to
Explain the Narrow Range of Diversity Among Species}

\author[1*]{Vince Buffalo}
\affil[1]{Institute of Ecology and Evolution, University of Oregon, United States}

\corr{vsbuffalo@gmail.com}{VB}

% \contrib[\authfn{1}]{These authors contributed equally to this work}
% \contrib[\authfn{2}]{These authors also contributed equally to this work}

% \presentadd[\authfn{3}]{Department, Institute, Country}
% \presentadd[\authfn{4}]{Department, Institute, Country}
% \presentadd[\authfn{5}]{eLife Sciences editorial Office, eLife Sciences, Cambridge, United Kingdom}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

@@

 Neutral theory predicts that genetic diversity increases with population size,
 yet observed levels of diversity across metazoans vary only two orders of
 magnitude while population sizes vary over several. This unexpectedly narrow
 range of diversity is known as Lewontin's Paradox of Variation (1974). While
 some have suggested selection constrains diversity, tests of this hypothesis
 seem to fall short. Here, I revisit Lewontin's Paradox to assess whether
 current models of linked selection are capable of reducing diversity to this
 extent. To quantify the discrepancy between pairwise diversity and census
 population sizes across species, I combine previously-published estimates of
 pairwise diversity from 172 metazoan taxa with estimates of census sizes.
 Using phylogenetic comparative methods, I show this relationship is
 significant accounting for phylogeny, but with high phylogenetic signal and
 evidence that some lineages experience shifts in the evolutionary rate of
 diversity deep in the past. Additionally, I find a negative relationship
 between recombination map length and census size, suggesting abundant species
 have less recombination and experience greater reductions in diversity due to
 linked selection. However, I show that even using strong selection parameter
 estimates, models of linked selection are unlikely to explain the observed
 relationship between diversity and census sizes across species.
\end{abstract}

\section{Introduction}

A longstanding mystery in evolutionary genetics is that the observed levels of
genetic variation across sexual species are confined to an unexpectedly narrow
range. Under neutral theory, the average number of nucleotide differences
between lineages (pairwise diversity, $\pi$) is determined by the balance of
new mutations and their loss by genetic drift
\citep{Kimura1964-ul,Malecot1948-zv,Wright1931-fl}. In particular, the
expected diversity at neutral sites in a panmictic population of $N_c$ diploids
is expected to be $\pi \approx 4N_c \mu$, where $\mu$ is the per basepair per
generation mutation rate. Given that metazoan germline mutation rates only
differ 10-fold ($10^{-8}$--$10^{-9}$, \cite{Kondrashov2010-fi,Lynch2010-ki}),
and census sizes vary over several orders of magnitude, one would expected
under neutral theory that heterozygosity should also vary over several orders
of magnitude. However, early allozyme surveys revealed that heterozygosity
levels across a wide range of species varied just an order of magnitude
\citep[][p.  208]{Lewontin1974-jb}; this is known as Lewontin's ``Paradox
of Variation".  With modern sequencing-based estimates of $\pi$ across taxa
ranging over only three orders of magnitude ($0.01$--$10\%$,
\cite{Leffler2012-zj}), Lewontin's paradox has persisted unresolved through the
genomics era.

From the beginning, explanations for Lewontin's Paradox have been framed in
terms of the neutralist--selectionist controversy
\citep{Lewontin1974-jb,Kimura1984-ia,Gillespie1991-qa,Gillespie2001-mv}.
The neutralist view is that beneficial alleles are sufficiently rare and
deleterious alleles removed sufficiently quickly, that levels of genetic
diversity are shaped predominantly by genetic drift and mutation
\citep{Kimura1984-ia}. Specifically, \emph{non-selective} processes
decouple the effective population size implied by observed levels of diversity
$\widehat{\pi}$, $\widetilde{N}_e = \nicefrac{\widehat{\pi}}{4\mu}$, from the
census size, $N_c$. By contrast, the selectionist view is that the direct and
indirect effects of linked selection suppress diversity levels across taxa,
specifically because the impact of linked selection is greater in large
populations. Undoubtedly, these opposing views represent a false dichotomy, as
population genomic studies have uncovered both complex demographic histories
that impact diversity within a species (e.g.
\cite{Zhao2013-vd,Palkopoulou2015-bg}) and evidence that selection depresses
genome-wide diversity (e.g.
\cite{Elyashiv2016-vt,Begun1992-ey,Aguade1989-jx,McVicker2009-ax}).

\subsection{Possible Explanations of Lewontin's Paradox}

A resolution of Lewontin's Paradox would involve a mechanistic description and
quantification of the evolutionary processes that prevent diversity from
scaling with census sizes across species. This would necessarily connect to the
broader literature on the empirical relationship between diversity and
population size
\citep{Frankham1996-yb,Nei1984-zi,Soule1976-he,Leroy2021-gy}, and the
ecological and life history correlates of genetic diversity
\citep{Nevo1978-wh,Powell1975-lg,Nevo1984-hp}. Three categories of
processes stand out as potentially capable of decoupling census sizes from
diversity: non-equilibrium demography, variance and skew in reproductive
success, and selective processes.

It has long been appreciated that effective population sizes are typically less
than census population sizes, tracing back to early debates between R.A. Fisher
and Sewall Wright \citep{Fisher1947-tf,Wright1948-fj}. Possible causes of
this divergence between effective and census population sizes include
demographic history (e.g. population bottlenecks), extinction and
recolonization dynamics, or the breeding structure of populations (e.g. the
variance in reproductive success and population substructure). Early
explanations for Lewontin's Paradox suggested bottlenecks during the last
glacial maximum severely reduced population sizes
\citep{Kimura1984-ia,Ohta1973-pk,Nei1984-zi}, and emphasized that large
populations recover to equilibrium diversity levels more slowly
(\cite{Nei1984-zi}, \cite{Kimura1984-ia} p. 203-204). Another explanation is
that cosmopolitan species repeatedly endure extinction and recolonization
events, which reduces effective population size
\citep{Maruyama1980-xz,Slatkin1977-kd}.

While chance demographic events like bottlenecks and recent expansions have
long-term impacts on diversity (since mutation-drift equilibrium is reached on
the order of size of the population), characteristics of the breeding structure
such as high variance ($V_w$) or skew in reproductive success continuously
suppress diversity below the levels predicted by the census size
\citep{Wright1938-tv}. For example, in many marine animals, females are
highly fecund, and dispersing larvae face extremely low survivorship, leading
to high variance in reproductive success
\citep{Waples2018-kb,Waples2013-wi,Hedgecock2011-ku,Hauser2008-fd}. Such
``sweepstakes" reproductive systems can lead to remarkably small ratios of
effective population size to census population size (e.g. $\nicefrac{N_e}{N_c}$
can range from $10^{-6}$--$10^{-2}$), since $\nicefrac{N_e}{N} \approx
\nicefrac{1}{V_w}$
\citep{Hedgecock1994-gs,Wright1938-tv,Nunney1993-ef,Nunney1996-wy}, and
require multiple-merger coalescent processes to describe their genealogies
\citep{Eldon2006-ui}. Overall, these reproductive systems diminish the
diversity in many species, but seem unlikely to explain Lewontin's Paradox
broadly across metazoans.

Alternatively, selective processes, and in particular the indirect effects of
selection on linked neutral variation, could explain the observed narrow range
of diversity. The earliest mathematical model of hitchhiking was proffered as a
explanation of Lewontin's Paradox \citep{Maynard_Smith1974-zr}.  Since,
empirical observations have demonstrated that linked selection shapes patterns
of genome-wide diversity, as evidenced by the correlation between recombination
and diversity in a variety of species
\citep{Aguade1989-jx,Begun1992-ey,Cutter2003-tl,Stephan1998-hh,Cai2009-by}.
Theoretic work to explain this pattern has considered diversity under a steady
influx of new beneficial mutations (recurrent hitchhiking;
\cite{Stephan1992-jc,Stephan1995-ry}), and purifying selection against new
deleterious mutations (background selection, BGS;
\cite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1994-oh,Hudson1995-xc}).
Indeed, empirical work indicates background selection diminishes diversity
around genic regions in a variety of species
\citep{McVicker2009-ax,Hernandez2011-gs,Charlesworth1996-px}, and now
efforts have shifted towards teasing apart the effects of positive and negative
selection on genomic diversity \citep{Elyashiv2016-vt}.

An important class of theoretic selection models pertaining to Lewontin's
Paradox are recurrent hitchhiking models that decouple diversity from the
census population size. These models predict diversity when strongly selected
beneficial mutations regularly enter and sweep through the population, trapping
lineages and forcing them to coalesce
\citep{Kaplan1989-sc,Gillespie2000-mh}. In general, decoupling occurs under
these hitchhiking models when the rate of coalescence due to selection is much
greater than the rate of neutral coalescence \citep[e.g.][equation
22]{Coop2012-cd}. Under other linked selection models, the resulting effective
population size is proportional to population size, and thus these models
cannot decouple diversity, all else equal. For example, models of background
selection and polygenic fitness variation predict diversity is proportional to
population size, mediated by the total recombination map length and the
deleterious mutation rate or fitness variation
\citep{Charlesworth1993-gb,Nicolaisen2012-vs,Nordborg1996-nq,Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}.

\subsection{Recent Approaches Towards Solving Lewontin's Paradox}

Recently, \cite{Corbett-Detig2015-gt} used population genomic data to
estimate the reduction in diversity due to background selection and hitchhiking
across 40 species, and showed the impact of selection increases with two
proxies of census population size, species range and with body size. They
argued this is evidence that selection could explain Lewontin's Paradox;
however, in a re-analysis, \cite{Coop2016-gx} demonstrated that the
observed scale of these reductions is insufficient to explain the
orders-of-magnitude shortfall between observed and expected levels of diversity
across species. Other recent work has found that certain life history
characteristics related to parental investment, such as propagule size, are
good predictors diversity in animals \citep{Romiguier2014-bp,Chen2017-nf}.
Nevertheless, while these diversity correlates are important clues, they do not
propose a mechanism by which these traits act to constrain diversity within a
few orders of magnitude.

@@

Past work looking at the relationship between $\pi$ and $N_c$ has been unable
to fully account for phylogenetic non-independence across taxa
\citep{Felsenstein1985-an}. To address this shortcoming, I use phylogenetic
comparative methods (PCMs) with a synthetic time-calibrated phylogeny to
account for shared phylogenetic history. Moreover, it is disputed whether
considering phylogenetic non-independence is necessary in population genetics,
since coalescent times are much less than divergence times
\citep{Whitney2010-ud,Lynch2011-qv}. Using PCMs, I address this by
estimating the degree of phylogenetic signal in the diversity census size
relationship, and investigating how these traits evolve along the phylogeny.

Finally, I explore whether the predicted reductions of diversity under
background selection and recurrent hitchhiking are sufficiently strong to
resolve Lewontin's Paradox. These predicted reductions in diversity across
species are estimated using strong selection parameters from \emph{Drosophila
melanogaster}, a species known to be strongly affected by linked selection.
Given the effects of linked selection are mediated by recombination map length,
I investigate how recombination map lengths vary with census population size
using data from a previously-published survey \citep{Stapley2017-fs}. I
find map lengths are typically shorter in large--census-size species,
increasing the effects of linked selection in these species, which might
further decouple diversity from census size. Still, I find the combined impact
of these selection models with available parameter estimates falls short in
explaining Lewontin's Paradox, and discuss future avenues through which the
Paradox of Variation could be fully resolved.

\section{Results}

\subsection{Estimates of Census Population Size}

\begin{figure}
  \includegraphics[width=0.7\linewidth]{figures/Nc_figure.pdf}
  \caption{The distribution of approximate census population sizes estimated by this
   study. Some phyla containing few species were excluded for clarity.}
   \label{fig:figure-nc}

   \figsupp[The relationship between body mass and population density found by
   \cite{Damuth1987-sg}, which is used to predict population densities.]{The
     source of this data is appendix table of \cite{Damuth1987-sg}; the color
     indicates Damuth's original group labels. The dashed line was estimated
     using a lognormal regression model in Stan. References to each measurement
   are available in \cite{Damuth1987-sg}.}{\includegraphics[]{figures/damuth.pdf}}\label{figsupp:sf1}

  \figsupp[The fraction of total species on earth included in this study's
  sample, per class.]{The color of the points represents phylum, and the size of
the point represents the absolute number of species by
class.}{\includegraphics{figures/species_counts.pdf}}\label{figsupp:species-counts}

\figsupp[Validation of this paper's range estimates procedure against the IUCN Red List's range estimates.]{
  The correspondence between the ranges estimated with the alpha hull
  method applied to GBIF data used in this paper and IUCN Red List's Extent
  of Occurrence for the subset of species in both datasets. Note that the
  IUCN Red List contains predominantly endangered species, which leads to
  ascertainment bias; still, the high correlation between the estimated
ranges shows the alpha hull method works well.}{\includegraphics[]{figures/iucn_redlist_ranges.pdf}}\label{figsupp:redlist-ranges}

\figsupp[Validation of this paper's range estimates against the categorical labels of \cite{Leffler2012-zj}.]{
  The estimated ranges using GBIF occurrence data, ordered within and
  colored by the original range category labels assigned in
\cite{Leffler2012-zj}.}{\includegraphics[]{figures/range_categories.pdf}}\label{figsupp:range-cats}

  \figsupp[The relationship between body length (meters) and body mass (grams)
  in the \cite{Romiguier2014-bp} data set.]{The relationship between body
    length (meters) and body mass (grams) in the \cite{Romiguier2014-bp} data
    set. This is used to infer body masses for taxa. The gray dashed line is the
  line of best fit inferred using
Stan.}{\includegraphics[]{figures/body_mass_length.pdf}}\label{figsupp:body-mass-length}

\end{figure}

A major impediment in quantifying Lewontin's Paradox has been estimating census
population sizes across many taxa, especially for extremely abundant,
cosmopolitan species that define the upper limit of ranges. Previous work has
surveyed the literature for census size estimates
\citep{Nei1984-zi,Soule1976-he,Frankham1996-yb}, or used range, body size, or
qualitative categories as proxies for census size
\citep{Corbett-Detig2015-gt,Leffler2012-zj}. To quantify the relationship
between genomic estimates of diversity and census population sizes, I first
approximate census population sizes for 172 metazoan taxa (\FIG{figure-nc}). My
approach predicts population densities from body sizes using a
previously-observed linear relationship that holds across metazoans (see
\FIGSUPP[figure-nc]{sf1}; \cite{Damuth1981-st,Damuth1987-sg}). Then, from
geographic occurrence data, I estimate range sizes.  Finally, I estimate
population size as the product of these predicted densities and range estimates
(see Methods and Materials: \nameref{sec:methods-popsize}). Note that the relationship between
population density and body size is driven by energy budgets, and thus reflects
macroecological equilibria \citep{Damuth1987-sg}. Consequently, population
sizes are underestimated for taxa like humans and their domesticated species,
and overestimated for species with anthropogenically reduced densities or
fragmented ranges. For example, the population size of \emph{Lynx lynx} is
likely around 50,000 \citep{Iucn2020-ap} which is around two orders of
magnitude smaller than my estimate. Additionally, the range size estimates do
not consider whether an area has unsuitable habitat, and thus may be
overestimated for species with particular niches or patchy habitats. While
these methods to estimate census size are crude and approximate, they can be
efficiently calculated for numerous taxa and are sufficient to estimate the
scale of Lewontin's Paradox (see \nameref{supp:popsize-valid} for more on
validation based on biomass and other approaches).

\subsection{Quantifying Lewontin's Paradox}

\begin{figure}
  \begin{fullwidth}
  \includegraphics[width=\linewidth]{figures/diversity_popsize_full.pdf}
